---
title: 'Class 3: Data Acquisition, Wrangling, Exploration'
author: "Nile"
date: "November 9, 2019"
output: html_document
---
` is called as acute or left quote

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```


```{r}
# Lets use this R chunk to load required libraries and set project working directory

# This library is used for formating text(bold, color fonnt, font size etc)
#install.packages("kableExtra")
library(kableExtra)

# Set working directory
setwd("E:/Eudreka/003-EDA/")

```

We will use IMDB Movies dataset which has data of about 5000 odd movies.  
Using this dataset we will solve some of the tasks given below to cover today's topic.  

1. `r text_spec("Know top 10 movies with the highest profit ?",color="blue")`
2. `r text_spec("Know top rated movies in the list and also average IMDB score ?",color="blue")`
3. `r text_spec("Plot a graphical representation to show number of movies  release each year ?",color="blue")`

## Phase 1 : DATA ACQUISITION
In this phase, based on the source from where we have to gather data, we write R code to acquire them.  

*So of the type of acquisition can be from:*

1. CSV(at rest)
2. CSV(from web) 
3. Excel
4. MySQL database
5. WebScraping (unstrcutred)
6. XML
7. JSON
8. API

Lets, now write command to acquire data from above source types  


```{r }
# Load CSV file from WEB
url = "https://data.baltimorecity.gov/api/views/dz54-2aru/rows.csv?accessType=DOWNLOAD"
download.file(url,destfile = "Cameras.csv" )

camera_data = read.csv("Cameras.csv",header=TRUE)
head(camera_data,10)
```


```{r }
# Load Excel File
#install.packages("readxl")
library(readxl)

movie_data = read_xlsx("movie_metadata_excel.xlsx",sheet = "movie_metadata")
head(movie_data,10)
tail(movie_data,10)
```


```{r }
# Load CSV file at rest
movie_data = read.csv("movie_metadata.csv",header=TRUE)
head(movie_data)
str(movie_data)

my_dataframe = as.data.frame(matrix(nrow = 3,ncol=4))
my_dataframe

```


```{r }
# # # Load from mySQL Database
# 
#install.packages("RMySQL")
 library(RMySQL)

 myDBConnection = dbConnect(MySQL(),user="genome",db="hg19",host="genome-mysql.cse.ucsc.edu")
# 
# # Get list of all tables
 allTableNames = dbListTables(myDBConnection)
 head(allTableNames)
# 
# #get records from certain tables
 dbGetQuery(myDBConnection,"select count(*) from affyCytoScan") #affyU133Plus2
 dbDisconnect(myDBConnection)
```


```{r}
# library(RJDBC)
# 
# library(RODBC)
# 
# #Database connection settings:
# driverClassName="com.Hana.jdbc.Driver"
# driverPath="mysql-connector-java-5.1.26.jar"
# dbConnectionString="jdbc:mysql://hostname:port/schemaname"
# userID="root"
# pwd="root"
# 
# #Importing the dataset
# driver=JDBC(driverClassName,driverPath)
```


```{r}
# Load data using webscrapping
#install.packages("XML")
#library(RCurl)
library(XML)
url = "http://news.bbc.co.uk/2/hi/uk_politics/8044207.stm"
#url_obj = getURL(url)
html_doc = htmlParse(url) 
html_table = readHTMLTable(html_doc)
html_table[[2]]

```

```{r}
# Load data from XML

#install.packages("xml2")
library(xml2)
xml_doc = read_xml("http://www.w3schools.com/xml/simple.xml")
rootNode = xml_root(xml_doc)
xml_name(rootNode)

```

#####   Heavy load, avoid executing !!!!!
```{r}
# Load data from JSON
#install.packages("RJSONIO")
library(RJSONIO)

json_data = fromJSON("https://data.ny.gov/api/views/9a8c-vfzj/rows.json?accessType=DOWNLOAD")

market_data = json_data[['data']]
market_data[[1]][[14]]
market_data[[2]][[14]]

```

We have now acquired the raw data, we will have to now make sure that data is clean and thus prepare it for model.  
  

## Phase 2 : DATA PREPARATION

This phase multiple sub-phases like :  

1. **Data Wrangling** : Here we do cleaning, standardize, derive or remove unwanted columns
  + Data subsetting
  + Data filtering
  + Removing null values
  + Text manipulation

2. **Data Exploration** : Here we explore and understand data
  + Find number of row and columns in dataset
  + Understanding and displaying intial top and last records of datset
  + Understanding and displaying names of columns
  + Understanding the structure of the dataset
  + Understading various categories/levels of a catagorical variable
  + Display all the records of dataset
  
3. **Data Visualization** : Here we do visual representation of data
  + Boxplot
  + Histogram
  + Scatter plot

  

## Data Wrangling
```{r}
# str(movie_data)
# direct_name = as.character(movie_data$director_name)
# head(direct_name)
# (direct_name)
# class(movie_data$color)

# Write R code for Data Wrangling

# Sampling of data
dim(movie_data)
names(movie_data)
movie_data[3,c("actor_3_name","actor_3_facebook_likes")]
movie_data[r,c]

color_movies = movie_data[movie_data[ ,"color"]=="Color",]
dim(color_movies)

```

```{r}
# Filtering of Data
# Filter only director_name and director_facebook_likes
names(movie_data)
onlyDirName_likes_data = movie_data[  ,c(2,5)] # OR
onlyDirName_likes_data = movie_data[,c("director_name","director_facebook_likes")]
dim(onlyDirName_likes_data)
head(onlyDirName_likes_data)

temp_1 = movie_data[,2:5]
names(temp_1)
dim(temp_1)


temp_2 = movie_data[2:5,2:5]
names(temp_2)
dim(temp_2)

```

```{r}
# Cleaning : Removing records which have NA(Blank)

# Identifying if dataset have NA 
anyNA(movie_data)

# Which cells have NA
head(is.na(movie_data))

# Identidy which columns have how many NA's
colSums(is.na(movie_data))

movie_data_clean = na.omit(movie_data)

anyNA(movie_data_clean)
dim(movie_data_clean)
dim(movie_data)
str(movie_data_clean)

```

```{r}
head(movie_data)

# Text Manipulation
head(movie_data_clean$movie_title)

# remove unwanted text from dataset
movie_data_clean$movie_title = sub(pattern = "Â",replacement = "",x = movie_data_clean$movie_title)

head(movie_data_clean$movie_title)
head(trimws(movie_data_clean$movie_title))

class(movie_data_clean$movie_title)


#movie_data_clean$movie_title = trimws(movie_data_clean$movie_title,which = "both")

```

## Data Exploration
```{r}
# Write R code for Data Exploration/Analysis
dim(movie_data_clean)

nrow(movie_data_clean)
ncol(movie_data_clean)

head(movie_data_clean)
tail(movie_data_clean)

names(movie_data_clean)

str(movie_data_clean)

levels(movie_data_clean$color)

summary(movie_data_clean)

#strptime(dataset_original$Date,format="%d-%b-%y")


```

# Lets work on Film Release date 

```{r}

class(movie_data_clean$release_date)
head(movie_data_clean$release_date)

movie_data_clean$release_date = strptime(movie_data_clean$release_date,format="%d-%m-%Y")

head(movie_data_clean$release_date)

library(lubridate)
movie_data_clean$start_month =   month(movie_data_clean$release_date)
movie_data_clean$start_day = day(movie_data_clean$release_date)
movie_data_clean$start_hour = hour(movie_data_clean$release_date)
movie_data_clean$start_wday = weekdays(movie_data_clean$release_date)

head(movie_data_clean)


```

We are in position to solve the 1st to problems now...  
1. `r text_spec("Know top 10 movies with the highest profit ?",color="blue")`  

```{r}
# We will take cleaned dataset
str(movie_data_clean)

movie_data_clean$profit = movie_data_clean$gross - movie_data_clean$budget
head(movie_data_clean$profit,10)

movie_data_clean = movie_data_clean[order(-movie_data_clean$profit),]

head(movie_data_clean$profit,10)

#cbind(movie_title,profit)

head(cbind(movie_data_clean$profit,movie_data_clean$movie_title),10)

```

2. `r text_spec("Know top rated movies in the list and also average IMDB score ?",color="blue")`
```{r}
# Top 10 movies my IMDB ratings/score
movie_data_clean = movie_data_clean[order(-movie_data_clean$imdb_score),]
head(cbind(movie_data_clean$movie_title,movie_data_clean$imdb_score),10)

# Average of imdb_score
mean(movie_data_clean$imdb_score)

```


## Data Visualization
  
```{r}
#Boxplot
boxplot(movie_data_clean$imdb_score,horizontal = TRUE)

```

```{r}
# Histogram

hist(movie_data_clean$title_year)

```
  
```{r}
# Scatter plot
plot(movie_data_clean$duration)

```

  
3. `r text_spec("Plot a graphical representation to show number of movies  release each year ?",color="blue")`  

```{r}
# Write R code for Data Visualization
# Histogram
max(movie_data_clean$title_year,na.rm = TRUE)
min(movie_data_clean$title_year,na.rm=TRUE)
m_breaks = seq(min(movie_data_clean$title_year,na.rm=TRUE),max(movie_data_clean$title_year,na.rm = TRUE), by=1)
hist(movie_data_clean$title_year,breaks=m_breaks)
hist_data = hist(movie_data_clean$title_year,breaks=m_breaks)

#Index of years where highest movies
#temp = hist_data[order(-hist_data$counts),]
which.max(hist_data$counts)
hist_data$mids[82]

```


```{r}

#
library(ggplot2)
ggplot(movie_data_clean,aes(x=title_year)) + geom_bar()
ggplot(movie_data_clean,aes(x=title_year)) + geom_bar()


library(dplyr)
# select title_year, count(*) FilmCOunt from movie_data_clean group by title_year order filmcount desc

NumberOfReleasesPerYear = movie_data_clean %>% select(title_year) %>%
                          group_by(title_year) %>%
                          summarise(FilmCount =n() ) %>% arrange(desc(FilmCount))

head(NumberOfReleasesPerYear)

```


