---
title: "Module 7 - Recommendation Engines"
author: "Vishal Desai"
date: "4/22/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

Module 7: Assignment

# - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

Perform the following activities in R

Task 1: Association Rule mining
  
   Load Adult data into R (Adult data is a built-in dataset in the arules package of R).
   Plot a graph of ten items with highest support.
   Generate association rules with lhs values as "capital-loss=None" and   
    "native-country=United-States"
   Use arulesViz package to generate some visualizations from patterns

# - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

INSTALL LIBRARIES

```{r INSTALL LIBRARIES}

# Load Adult data into R (Adult data is a built-in dataset in the arules package of R).

install.packages("arules")
library(arules)
library(help=arules)
data("Adult")
adlts = Adult

```


DATA EXPLORATION

```{r DATA EXPLORATION}

summary(adlts)
View(adlts)
class(adlts)
str(adlts)
names(adlts)
nrow(adlts)
ncol(adlts)
colSums(is.na(adlts))

```

# - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

# Given number of ROWS and COLUMNS of our Dataset and Density shown how much of the Sparse matrix is filled...  <<  48842 * 115 * 0.1089939 = 612200.2  >>
# This are number of cell of our sparse matrix having values.

# - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

# This are the MOST FREQUENT ITEMS...

# most frequent items:
#           capital-loss=None            capital-gain=None native-country=United-States 
#                       46560                        44807                        43832 
#                  race=White            workclass=Private                      (Other) 
#                       41762                        33906                       401333 

# - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

# Gives the number of Transactions with given number of items.
# Eg. There are 19 Transactions with 9 Items and 30162 Transactions with 13 Items.

# element (itemset/transaction) length distribution:
# sizes
#    9    10    11    12    13 
#   19   971  2067 15623 30162 

# - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

```{r }

# Inspect function is used to view the association rules
inspect(adlts[1:3])

# Item frequency helps in getting support characteristics
itemFrequency(adlts[,1])
itemFrequency(adlts[,1:100])

# PLOT used to find out Support Values of all Items in Dataset
itemFrequencyPlot(adlts, support=0.08)

# - - - - - - - - - - - - - - - - -

# TASK: Plot a graph of ten items with highest support.

# PLOT used to find out Support Values of TOP 10 Items in Dataset
itemFrequencyPlot(adlts, topN =10)

# - - - - - - - - - - - - - - - - -

```

```{r }

# Generating association rules

m1 <- apriori(adlts)
summary(m1)  # set of 6137 rules
itemFrequencyPlot(adlts, support=0.08)

m2 <- apriori(adlts, parameter = list(support=0.097, confidence = 0.97))
summary(m2)
itemFrequencyPlot(adlts, support=0.097)

m3 <- apriori(adlts, parameter = list(support=0.099, confidence = 0.99))
summary(m3)
itemFrequencyPlot(adlts, support=0.099)

inspect(m3[1:10])

# Sort Rules where LIFT is HIGH
inspect(sort(m3, by = "lift")[1:5])

```


```{r }

# Generate association rules with lhs values as "capital-loss=None" and "native-country=United-States"

m4 = apriori(adlts, parameter = list(support=0.085,confidence=0.085),
             appearance = list(default="rhs", lhs = c("capital-loss=None","native-country=United-States")))
summary(m4)
inspect(m4[1:5])

install.packages("arulesViz")
library(arulesViz)
plot(m4)


```

```{r }

install.packages("arulesViz")
library(arulesViz)
plot(m4)

```

