---
title: "005 CLASSIFICATION - ASSIGNMENT"
author: "Vishal Desai"
date: "4/13/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Module 5: Assignment

Analyze the information given in the Affairs dataset and create classifiers using it.

The dataset can be loaded into R using the library:

➢ library(AER)

# IMPORT Library and required Dataser from the Library

```{r Import Library, warning=FALSE}

library(AER)
library(help=AER)
data("Affairs")
affairs = Affairs
summary(affairs)
str(affairs)
print(ncol(affairs))
print(nrow(affairs))

```

The description of the attributes in the dataset are as follows:

• affairs        –  Numeric Variable   -   How often engaged in extramarital sexual intercourse during the past year.

• gender         –  Factor             -   Indicating Gender.

• age            –  Numeric Variable   -   Coding age in years

• yearsmarried   –  Numeric Variable   -   Coding number of years married...
                                           0.125 =  3 months or less,
                                           0.417 =  4–6 months,
                                           0.75  =  6 months – 1 year,
                                           1.5   =  1–2 years,
                                           4     =  3–5 years,
                                           7     =  6–8 years,
                                           10    =  9–11 years,
                                           15    =  12 or more years.

• religiousness  –  Numeric Variable   -   Coding Religiousness:
                                           1  =  anti,
                                           2  =  not at all,
                                           3  =  slightly,
                                           4  =  somewhat,
                                           5  =  very.

• Education      -  Numeric Variable   -   coding level of education:
                                           9   =  Grade School,
                                           12  =  High School Graduate,
                                           14  =  Some College,
                                           16  =  College Graduate,
                                           17  =  Some Graduate Work,
                                           18  =  Master's Degree,
                                           20  =  Ph.D., M.D., or other advanced degree.

```{r}
hist(affairs$affairs)
```

Above Histrogram shows that number of Affairs individual having but most of the individual range from 0 -1 Affair

BRIEF INSIGHT:

From below extraction we can get little insight about the data, which are as follow...

There are 451 individuals who being in affair didn't had any physical relationship.
```{r}

affairs_aggregate = aggregate(x = affairs, by = list(unique.values = affairs$affairs), FUN=length)
affairs_aggregate

```

Female were highly involved in Physical relationship as compare of males, with figure showing 315 FEMALES and 286 MALES

```{r}
gender_aggregate = aggregate(x = affairs, by = list(unique.values = affairs$gender), FUN=length)
gender_aggregate
```

Three Age group had been involved highly in Physical Relationship which falls near the age of marriage, that being 22, 27 32.

And Three Age group who were less likely involved are 17.5, 52 & 57.

```{r}
aggregate(x = affairs, by = list(unique.values = affairs$age), FUN=length)
```

```{r}

aggregate(x = affairs, by = list(unique.values = affairs$yearsmarried), FUN=length)

```

• yearsmarried   –  Numeric Variable   -   Coding number of years married...
                                           0.125 =  3 months or less,
                                           0.417 =  4–6 months,
                                           0.75  =  6 months – 1 year,
                                           1.5   =  1–2 years,
                                           4     =  3–5 years,
                                           7     =  6–8 years,
                                           10    =  9–11 years,
                                           15    =  12 or more years.

Individual who were married and were highly involved in Physical Relationship were in range of 1. 12 years and more in a marriage, 3-5 years individuals and 1-2 years. 

```{r}

aggregate(x = affairs, by = list(unique.values = affairs$children), FUN=length)

```

Individuals with Children were highly involed in Physical Relationship that is 430.

```{r}

aggregate(x = affairs, by = list(unique.values = affairs$religiousness), FUN=length)

```

• religiousness  –  Numeric Variable   -   Coding Religiousness:
                                           1  =  anti,
                                           2  =  not at all,
                                           3  =  slightly,
                                           4  =  somewhat,
                                           5  =  very.

People who where somewhat religious were involved highly in Physical Relationship, range being Somewhat - 190, Not at all - 164 and Slightly - 129.

```{r}

aggregate(x = affairs, by = list(unique.values = affairs$education), FUN=length)

```

• Education      -  Numeric Variable   -   coding level of education:
                                           9   =  Grade School,
                                           12  =  High School Graduate,
                                           14  =  Some College,
                                           16  =  College Graduate,
                                           17  =  Some Graduate Work,
                                           18  =  Master's Degree,
                                           20  =  Ph.D., M.D., or other advanced degree.

Individuals who's education level was just till High School were less likely involved in Physical Relationship as compare to who did completed Graduation and Master's Degree, figure being...

Some COllege - 154, College Graduates - 115 and Master's - 112 and on the least side were individuals who studided only till Grade School - 7 and individuals who had High School were also less likely involved number being to 44.


```{r}

aggregate(x = affairs, by = list(unique.values = affairs$occupation), FUN=length)

```

Look for NA Values in our Dataset...

```{r}
colSums(is.na(affairs))
```

Task 1: Using the affairs column of our data

   Create a new column with nominal values “YES” and “NO”.
  
   Convert it into factor



```{r}
# Create a copy of our Dataset
aff_new = affairs

# Have a look at new Dataset
str(aff_new)

```

```{r}

# Create a New Column
aff_new$aff_yes_no = aff_new$affairs

# Replace the 0 with No and all other numbers to Yes...
aff_new$aff_yes_no = sub("0", "No", aff_new$aff_yes_no)
aff_new$aff_yes_no = sub("12", "Yes", aff_new$aff_yes_no)
aff_new$aff_yes_no = sub("1", "Yes", aff_new$aff_yes_no)
aff_new$aff_yes_no = sub("2", "Yes", aff_new$aff_yes_no)
aff_new$aff_yes_no = sub("3", "Yes", aff_new$aff_yes_no)
aff_new$aff_yes_no = sub("7", "Yes", aff_new$aff_yes_no)
view(aff_new$aff_yes_no)

```

```{r}

# Check Class of new created columns after replacing the number to Yes and No
str(aff_new)
class(aff_new$aff_yes_no)

# Change the Class of new columns to Factor
aff_new$aff_yes_no = as.factor(as.factor(aff_new$aff_yes_no))
view(aff_new$aff_yes_no)

```

```{r Export CSV File}

# Save our New file for usage and avoid any mishap
write.csv(aff_new, "Edited_Affairs_DataSet.csv")

```

Task 2: Create a classifier with our data using Decision Tree Algorithm

   Plot the Decision tree
  
   Calculate the accuracy using confusion matrix


#####################
  DECISION TREE
#####################
```{r DECISION TREE}

#Import Dataset
affairs_new = read.csv("Edited_Affairs_DataSet.csv")
str(affairs_new)
summary(affairs_new)
affairs_new = affairs_new[, -1]
View(affairs_new)
affairs_new$affairs = affairs_new$aff_yes_no
View(affairs_new)
affairs_new = affairs_new[, -10]
View(affairs_new)


#Install Library
install.packages("rpart")
library(rpart)

#Build a training and testing set
set.seed(4)
id = sample(2, nrow(affairs_new), prob = c(0.7,0.3),replace = T)
trainset = affairs_new[id==1,]
testset = affairs_new[id==2,]
class(trainset)

View(trainset)
View(testset)

# Build decision tree model
d_tree_model = rpart(trainset$affairs~. ,data = trainset)
summary(d_tree_model)

### Depth of our DECISION TREE is     =  419
### Important Variable being          =  RATING
### Least Important Variable being    =  RELIGIOUSNES


plot(d_tree_model, margin=0.1)
text(d_tree_model, use.n = TRUE, pretty = TRUE, cex=0.8)

#temp = trainset[trainset$glucose_conc<154.5 & trainset$BMI<26.35,]
#table(temp$Is_Diabetic)

predvalues_new = predict(d_tree_model, newdata = testset, type = "class")
predvalues_new

View(predvalues_new)
View(affairs_new$affairs)
str((predvalues_new))
str(affairs_new$affairs)

install.packages("caret")
library(caret)

install.packages("e1071")
library(e1071)
confusionMatrix(table(predvalues_new, testset$affairs))

```


#####################
  
  RANDOM FOREST
  
#####################

```{r IMPORT DATASER - RANDOM FOREST}

getwd()
df_affairs = read.csv("Edited_Affairs_DataSet.csv")
str(df_affairs)
df_affairs = df_affairs[, -1]
df_affairs$affairs = df_affairs$aff_yes_no
df_affairs = df_affairs[, -10]

```

```{r Import Library}
install.packages("randomForest")
library(randomForest)

```

```{r TRAIN-TEST SPLIT}

#Build a training and testing set
set.seed(4)
id = sample(2,nrow(df_affairs), prob = c(0.7,0.3), replace = T)
trainset = df_affairs[id == 1,]
testset = df_affairs[id==2,]
class(trainset)

#Build random forest model
rd_frst_model = randomForest(trainset$affairs ~ ., data = trainset, ntree=200)
rd_frst_model
plot(rd_frst_model)

####
# No. of variables tried at each split: 2
# OOB estimate of  error rate: 26.97%


```

```{r}

library(data.table)
library(ggplot2)

# Get OOB data from plot and coerce to data.table
oobData = as.data.table(plot(rd_frst_model))

# Define trees as 1 : ntree
oobData[, trees := .I]

# Cast to long format
oobData2 = melt(oobData, id.vars = "trees")
setnames(oobData2, "value", "error")

# Plot using ggplot
ggplot(data = oobData2, aes(x = trees, y = error, color = variable)) + geom_line()

predvalues<-predict(rd_frst_model, newdata = testset, type = "class")
predvalues

```

# CONFUSION MATRIX

```{r}

install.packages("caret")
library(caret)

install.packages("e1071")
library(e1071)
confusionMatrix(table(predvalues, testset$affairs))

# Confusion Matrix and Statistics = 78.57

```

```{r}

hist(treesize(rd_frst_model))

varImp(rd_frst_model)
varImpPlot(rd_frst_model, sort = TRUE, type = 2)
importance(rd_frst_model)
varUsed(rd_frst_model)

```

```{r}

# Using For loop to identify the right mtry for model
a = c()
i = 5
for (i in 3:8) {
  model3 <- randomForest(trainset$affairs ~ . , data = trainset, ntree = 500, mtry = i, importance = TRUE)
  predValid <- predict(model3, testset, type = "class")
  a[i-2] = mean(predValid == testset$affairs)
}

a

plot(3:8,a)


```